Davos: a Python package ``smuggler'' for constructing lightweight reproducible notebooks

### 2023-09-29 paper/main.tex
some additional wording updates to section 3

Lines 30--37 account for both of these possibilities programmatically:
\begin{center}
\includegraphics[width=0.9\textwidth]{figs/example7}
\end{center}
First, in line 30, we set the \texttt{davos.project} attribtue to \texttt{None} to temporarily 
%disable Davos's project isolation system and
allow installing smuggled packages outside of the notebook's project directory. 
As noted in Section \ref{subsec:projects}, this is typically discouraged, as doing so can risk interfering with the user's Python environment if existing package versions are overwritten.
In this particular case, however, a combination of factors make this relatively safe and inconsequential.
First, the package we need to install directly into the notebook server environment (\texttt{widgetsnbextension}) is smuggled without an accompanying onion comment (line 34), meaning that Davos will not replace any version the user may already have installed.
Second, the package has no dependencies of its own, so if Davos does install it, no other packages could potentially be installed or updated as a side effect.
Third, the package itself provides no functionality outside of rendering Jupyter widgets, so its presence would not alter any other code's expected behavior.

Next, in lines 31--33, we change the \texttt{pip} executable Davos uses to install smuggled packages (see Sec.~\ref{subsec:config}), storing the default executable's path in a variable before doing so. 
When Davos's project system is disabled, using a \texttt{pip} executable from a particular Python environment will cause smuggled packages to be installed into (and subsequently loaded from) that environment. 
The default \texttt{pip\_executable} will install packages into the environment used to run the IPython kernel. 
Here, the new value assigned to \texttt{davos.pip\_executable} in line 33 is the output of running ``\texttt{command -v pip}'' as a \texttt{!}-prefixed IPython system shell command in line 32 (``\texttt{command -v}'' outputs the path to an executable, similar to ``\texttt{which}'' but more portable). 
Since IPython system shell command are always executed in the notebook server environment, this command's output will be the path to that environment's \texttt{pip} executable---which may or may not be different from the kernel environment's.

After smuggling the \texttt{widgetsnbextension} package in line 34, we use the \texttt{davos.use\_default\_project()} function in line 35 to revert to installing package into the notebook's project directory, restore the default value of \texttt{davos.pip\_executable} in line 36, and \texttt{smuggle} the specified version of \texttt{ipywidgets} in line 37.
With these two packages now installed
and imported, line 39 smuggles \texttt{tqdm}~\cite{daCoEtal22}, which
displays progress bars to provide status updates for running code. In
Jupyter notebooks, the \texttt{tqdm.notebook} module can be imported
to enable more aesthetically pleasing progress bars that are displayed via
\texttt{ipywidgets}, if that package is installed and
importable. Therefore, to take advantage of this feature, it was
important to \texttt{smuggle} \texttt{tqdm} after ensuring the
\texttt{ipywidgets} package was available.

\stoppedhere

Next, we load in the reformatted dataset (line 33) and pre-trained
model (line 35) that we wish to use in our analysis.  In our
hypothetical example, we can suppose that the model was provided as a
\texttt{scikit-learn} \texttt{Pipeline} object that passes data
through two pretrained models in succession.  First, a trained \texttt{CountVectorizer}
instance converts text data to an array of word counts.  Second, the
word counts are passed to a topic model~\cite{BleiEtal03} using a
pretrained \texttt{LatentDirichletAllocation} instance.
\begin{center}
\includegraphics[width=0.9\textwidth]{figs/example8}
\end{center}
Let us suppose that the \texttt{Pipeline} object had been saved by its
original creator using the \texttt{joblib} package, as
\texttt{scikit-learn}'s documentation recommends \cite{skle22}.  Because
\texttt{joblib} uses the \texttt{pickle} protocol internally, the
ability to save and load pre-trained models is not guaranteed across
different \texttt{scikit-learn} versions.  For example, suppose that
the \texttt{Pipeline} object was created using \texttt{scikit-learn}
v0.21.3.  In that version of \texttt{scikit-learn}, the
\texttt{LatentDirichletAllocation} class was defined in
\texttt{sklearn.decomposition.online\_lda}.  However, in version
0.22.0, that module was renamed to \texttt{\_online\_lda}, and in
version 0.22.1, it was again renamed to \texttt{\_lda}.

>>>>

Lines 30--37 account for both of these possibilities programmatically:
\begin{center}
\includegraphics[width=0.9\textwidth]{figs/example7}
\end{center}
First, in line 30, we set the \texttt{davos.project} attribute to \texttt{None} to temporarily 
%disable Davos's project isolation system and
allow installing smuggled packages outside of the notebook's project directory. 
As noted in Section \ref{subsec:projects}, this is typically discouraged, as doing so can risk interfering with the user's Python environment if existing package versions are overwritten.
In this particular case, however, a combination of factors make this relatively safe and inconsequential.
First, the package we need to install directly into the notebook server environment (\texttt{widgetsnbextension}) is smuggled without an accompanying onion comment (line 34), meaning that Davos will not replace any version the user may already have installed.
Second, the package has no dependencies of its own, so if Davos does install it, no other packages could potentially be installed or updated as a side effect.
Third, the package itself provides no functionality outside of rendering Jupyter widgets, so its presence would not alter any other code's expected behavior.

Next, in lines 31--33, we change the \texttt{pip} executable Davos uses to install smuggled packages (see Sec.~\ref{subsec:config}), storing the default executable's path in a variable before doing so. 
When Davos's project system is disabled, using a \texttt{pip} executable from a particular Python environment will cause smuggled packages to be installed into (and subsequently loaded from) that environment. 
The default \texttt{pip\_executable} will install packages into the environment used to run the IPython kernel. 
Here, the new value assigned to \texttt{davos.pip\_executable} in line 33 is the output of running ``\texttt{command -v pip}'' as a \mbox{\texttt{!}-prefixed} IPython system shell command in line 32 (``\texttt{command -v}'' outputs the path to an executable, similar to ``\texttt{which}'' but more portable). 
IPython system shell command are always executed in the notebook server's environment---which may or may not be different from the kernel's environment---so this command's output will be the path to the server environment's \texttt{pip} executable.

After smuggling the \texttt{widgetsnbextension} package in line 34, we use the \texttt{davos.use\_default\_project()} function in line 35 to revert to installing package into the notebook's project directory, restore the default value of \texttt{davos.pip\_executable} in line 36, and \texttt{smuggle} the specified version of \texttt{ipywidgets} in line 37.
With these two packages now installed
and imported, line 39 smuggles \texttt{tqdm}~\cite{daCoEtal22}, which
displays progress bars to provide status updates for running code. In
Jupyter notebooks, the \texttt{tqdm.notebook} module can be imported
to enable more aesthetically pleasing progress bars that are displayed via
\texttt{ipywidgets}, if that package is installed and
importable. Therefore, to take advantage of this feature, it was
important to first ensure ensure that both \texttt{ipywidgets} and \texttt{widgetsnbextension} were available.

Next, we load in the reformatted dataset (line 33) and pre-trained
model (line 35) that we wish to use in our analysis.  In our
hypothetical example, we can suppose that the model was provided as a
\texttt{scikit-learn} \texttt{Pipeline} object that passes data
through two pretrained models in succession.  First, a trained \texttt{CountVectorizer}
instance converts text data to an array of word counts.  The
word counts are then passed to a topic model~\cite{BleiEtal03} using a
pretrained \texttt{LatentDirichletAllocation} instance.
\begin{center}
\includegraphics[width=0.9\textwidth]{figs/example8}
\end{center}
Let us suppose that the \texttt{Pipeline} object had been saved by its
original creator using the \texttt{joblib} package, as
\texttt{scikit-learn}'s documentation recommends \cite{skle22}.  Because
\texttt{joblib} uses the \texttt{pickle} protocol internally, the
ability to save and load pre-trained models is not guaranteed across
different \texttt{scikit-learn} versions.  For example, suppose that
the \texttt{Pipeline} object was created using \texttt{scikit-learn}
v0.21.3.  In that version of \texttt{scikit-learn}, the
\texttt{LatentDirichletAllocation} class was defined in
the \texttt{sklearn.decomposition.online\_lda} module.  However, in version
0.22.0, that module was renamed to \texttt{\_online\_lda}, and in
version 0.22.1, it was again renamed to \texttt{\_lda}.
